Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-12 paper/disc20clustering.tex
font issues with \OPT \RM

 \newcommand{\OPT}{\textsc{OPT}\xspace}
  \newcommand{\OPTM}{\mathit{OPT}}
 \newcommand{\ALG}{\textsc{ALG}\xspace}
 \newcommand{\PPL}{\textsc{PPL}\xspace}
 \newcommand{\OBRP}{BRP\xspace}
 \newcommand{\PPOBRP}{PP-BRP}
 \newcommand{\dist}{\mathit{dist}}
 \newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace}
  \newcommand{\RM}{\textsc{RM}\xspace} % rematching alg

@@

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	Fix an input sequence of requests
$\sigma:=\{\sigma_1,\dots, \sigma_m\}$.
	Assume on request $\sigma_{t}=\set{u,v}$, 
	i.e., at (time) $t$,
	\RM collocates the pair $\set{u,v}$.
	This means,
	the value of $C_{\set{u,v}}$ at $t$,
	denoted by $C^t_{\set{u,v}}$, 
	reaches $\lambda$ immediately before \RM resets the counter,
	i.e.,
	$ C^t_{\set{u,v}} = \lambda$.
	We denote the set of all requests to a~pair $\set{x,y}$ that arrive
	during $[t_1,t_2]$ by $\sigma_{\set{x,y}}[t_1,t_2]$,
	and its overall cost to \OPT by
	$\mathit{OPT} (\sigma_{\{u,v\}}[t_1,t_2])$.
	We may use $\sigma_{\set{x,y}}$ whenever
	 the interval $[t_1,t_2]$ is clear from the context.
	%
	%Else, this is not the first time and
	%Alg collocates $\{u,v\}$ also some time prior to $t$.

@@

	First,
	we bound  costs incurred by \RM for requests that
	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
	is the set of times when \RM performs a~collocation.
	By definitions of $t$ and $t'$,
	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by \RM,
	i.e., the total cost of remote serving 	and the moving cost is
	$\lambda + 2\alpha$.	
	Next,
	we bound costs incurred by \RM
	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
	Assume $\{u,v\}$ is not collocated at $t=m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means \RM pays $C^{m}_{ \{u,v\} }$
	for  requests in $\sigma_{\set{u,v}}(t',m]$.
    Then the overall cost to \RM is
$	\mathit{RM} (\sigma)
=
\sum_{ t \in T}(\lambda + 2\alpha) +
\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
$.

	Next,
	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
	If $t$ is the first time that \RM collocates $\{u,v\}$,
	 then  \OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with  cost zero.
	Therefore in this case,
	$\OPTM (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	Otherwise,
	it is not the first collocation and
	consider times $t'$ and $t''$ as define previously,
	 and let 
	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
	We define $R_{t'}$ for the collocation at $t'$  analogously.
	Then,
	$\OPTM(R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
	+ \OPTM(\sigma_{\{u,v\}}) $.
	If \OPT has both pairs separated during their respective intervals,
	then obviously it pays $2\lambda$ during those intervals.
	Note that \OPT cannot have both pairs collocated at the same time.
%	However, it can have one of the pairs already collocated prior to its respective interval.
	Let us assume \OPT has one of the pairs,
	e.g.~$\{u,v\}$,
	 collocated already prior its respective interval, $(t',t]$,
	 and keeps it so during the interval.
	 Then it pays zero while serving $\sigma_{\{u,v\}}$.
	Hence,
	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
	Therefore in any case,
	$\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.

	It remains to bound the cost  incurred by \OPT due to requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
	We bound the cost analogously to the case where \RM collocates $\{u,v\}$.
	If $\{u,v\}$ is not collocated in the initial matching
	and \RM never collocates it,
	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
	\OPT pays
	$\mathit{OPT} (\sigma_{\{u,v\}}[1,m]) 
	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
	for collocating this pair or (and) paying for (resp.~some of) requests in $\sigma_{\{u,v\}}[1,m]$.
	Else,
	either $\{u,v\}$ is collocated in the initial matching
	or \RM collocates it at some point.
	Then there exists an intersecting pair $\set{u,w}$
	that is collocated by \RM at $t' < m$,
	separating $\{u,v\}$.
	We define times $t'' < t' < m$ similarly to the former case.
	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
	Then,
	\OPT must pay for collocating at least one pair or (and) serving requests 
	to the other pair remotely.
	Thus,
	$\mathit{OPT} (R^*_{\{u,v\}}) 
	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.

	Next, we sum up all costs incurred by \OPT.
	By definitions of $R_t$ and $R^*_{\{u,v\}}$, we have either
	$R_{t'} \cap R_t = \sigma_{\{u,w\}}$ or
	$R_{t'} \cap R^*_{\{u,v\}} = \sigma_{\{u,w\}}$
	(\mahmoud{See figure \ref{fig:doubleCount}}). 
	This means,
	$\OPTM ( \sigma_{\{u,w\}})$
	is counted at most twice in each of  the expressions
	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R_t)$
	and  
	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R^*_{\{u,v\}})$.
	Hence,
	for all collocations performed by \RM,
	and for final requests at $t=m$,
	\OPT pays at least 
	$\frac{1}{2}(
	\sum_{ t \in T } \mathit{OPT} (R_t) +
	\sum_{\{u,v\}} \mathit{OPT} (R^*_{\{u,v\}})
	) $.
	Then,
	the total cost to \OPT is

	\begin{align*} 	%\label{eq:costOPT}
		\mathit{OPT} (\sigma)
		&=
		\frac{1}{2}
		\Big(
		\sum_{ t \in T} \mathit{OPT} (R_t) 
		+ \sum_{\{u,v\}}\mathit{OPT} (R^*_{\{u,v\}})
		\Big)	\\
		&\geq
		\Big(
		\sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
		\sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
		\Big)		
		=
		\frac{1}{2}		
		\Big(
		\sum_{ t \in T} \alpha  
		+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}
		\Big),
	\end{align*}
and
$
	\mathit{ALG} (\sigma)	/
	\mathit{OPT} (\sigma)
	\leq
	2\Big(
	\sum_{ t \in T} 3\alpha +
	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
	\Big)	 \big /
	\Big(
	\sum_{ t \in T} \alpha  
	+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}  
	\Big) 	\leq 6.
$
\end{proof}

@@

\begin{algorithm}
	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
	\renewcommand{\algorithmicensure}{\textbf{Output:}}
	\begin{algorithmic}
		%        \Require 
		%        $k, \ell$,
		%        initial partition $P_I$,
		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
		%        \Ensure a~final partition $P_F$ 
		\STATE {For each node $v$ create a~singleton component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
		%\STATE{$P := P_I$}
		\label{line:initcomponents}
		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
		\STATE{Let $C_1 \ni u$ and $C_2 \ni v$ be the  components containing $u$ and $v$, respectively.}
		\IF{$C_1 \neq C_2$}
		\STATE {Merge $C_1$ and $C_2$ into one component $C'$ and
			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
		%		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
		%		\COMMENT{i.e.~if not in the same cluster.}}       
		\IF{$C_1$ and $C_2$ are not collocated}
		\STATE {Move to a~partition closest to $P_I$ and respecting all components in $\mathcal{C}$.} 
		%		\COMMENT{move to a~partition closest to $P_I$}
		\label{line:rebalance} 
		\ENDIF
		\ENDIF
		\ENDFOR
	\end{algorithmic}
	\caption{Perfect Partition Learner (\PPL)}
	\label{alg:ppl}
\end{algorithm}

>>>>

 \newcommand{\OPT}{\textsc{OPT}\xspace}
 \newcommand{\ALG}{\textsc{ALG}\xspace}
 \newcommand{\PPL}{\textsc{PPL}\xspace}
 \newcommand{\OBRP}{BRP\xspace}
 \newcommand{\PPOBRP}{PP-BRP}
 \newcommand{\dist}{\textsc{dist}}
 \newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace}
  \newcommand{\RM}{\textsc{RM}\xspace} % rematching alg

@@

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	Fix an input sequence of requests
$\sigma:=\{\sigma_1,\dots, \sigma_m\}$.
	Assume on request $\sigma_{t}=\set{u,v}$, 
	i.e., at (time) $t$,
	\RM collocates the pair $\set{u,v}$.
	This means,
	the value of $C_{\set{u,v}}$ at $t$,
	denoted by $C^t_{\set{u,v}}$, 
	reaches $\lambda$ immediately before \RM resets the counter,
	i.e.,
	$ C^t_{\set{u,v}} = \lambda$.
	We denote the set of all requests to a~pair $\set{x,y}$ that arrive
	during $[t_1,t_2]$ by $\sigma_{\set{x,y}}[t_1,t_2]$,
	and its overall cost to \OPT by
	$\OPT (\sigma_{\{u,v\}}[t_1,t_2])$.
	We may use $\sigma_{\set{x,y}}$ whenever
	 the interval $[t_1,t_2]$ is clear from the context.
	%
	%Else, this is not the first time and
	%Alg collocates $\{u,v\}$ also some time prior to $t$.

@@

	First,
	we bound  costs incurred by \RM for requests that
	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
	is the set of times when \RM performs a~collocation.
	By definitions of $t$ and $t'$,
	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by \RM,
	i.e., the total cost of remote serving 	and the moving cost is
	$\lambda + 2\alpha$.	
	Next,
	we bound costs incurred by \RM
	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
	Assume $\{u,v\}$ is not collocated at $t=m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means \RM pays $C^{m}_{ \{u,v\} }$
	for  requests in $\sigma_{\set{u,v}}(t',m]$.
    Then the overall cost to \RM is
$	\RM (\sigma)
=
\sum_{ t \in T}(\lambda + 2\alpha) +
\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
$.

	Next,
	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
	If $t$ is the first time that \RM collocates $\{u,v\}$,
	 then  \OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with  cost zero.
	Therefore in this case,
	$\OPT(\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	Otherwise,
	it is not the first collocation and
	consider times $t'$ and $t''$ as define previously,
	 and let 
	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
	We define $R_{t'}$ for the collocation at $t'$  analogously.
	Then,
	$\OPT(R_t) =\OPT (\sigma_{\{u,w\}}) 
	+ \OPT(\sigma_{\{u,v\}}) $.
	If \OPT has both pairs separated during their respective intervals,
	then obviously it pays $2\lambda$ during those intervals.
	Note that \OPT cannot have both pairs collocated at the same time.
%	However, it can have one of the pairs already collocated prior to its respective interval.
	Let us assume \OPT has one of the pairs,
	e.g.~$\{u,v\}$,
	 collocated already prior its respective interval, $(t',t]$,
	 and keeps it so during the interval.
	 Then it pays zero while serving $\sigma_{\{u,v\}}$.
	Hence,
	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
	Therefore in any case,
	$\OPT(R_t) \geq  \min{\{ \lambda,\alpha \}}$.

	It remains to bound the cost  incurred by \OPT due to requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
	We bound the cost analogously to the case where \RM collocates $\{u,v\}$.
	If $\{u,v\}$ is not collocated in the initial matching
	and \RM never collocates it,
	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
	\OPT pays
	$\OPT(\sigma_{\{u,v\}}[1,m]) 
	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
	for collocating this pair or (and) paying for (resp.~some of) requests in $\sigma_{\{u,v\}}[1,m]$.
	Else,
	either $\{u,v\}$ is collocated in the initial matching
	or \RM collocates it at some point.
	Then there exists an intersecting pair $\set{u,w}$
	that is collocated by \RM at $t' < m$,
	separating $\{u,v\}$.
	We define times $t'' < t' < m$ similarly to the former case.
	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
	Then,
	\OPT must pay for collocating at least one pair or (and) serving requests 
	to the other pair remotely.
	Thus,
	$\OPT (R^*_{\{u,v\}}) 
	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.

	Next, we sum up all costs incurred by \OPT.
	By definitions of $R_t$ and $R^*_{\{u,v\}}$, we have either
	$R_{t'} \cap R_t = \sigma_{\{u,w\}}$ or
	$R_{t'} \cap R^*_{\{u,v\}} = \sigma_{\{u,w\}}$
	(\mahmoud{See figure \ref{fig:doubleCount}}). 
	This means,
	$\OPT ( \sigma_{\{u,w\}})$
	is counted at most twice in each of  the expressions
	$\OPT (R_{t'}) + \OPT (R_t)$
	and  
	$\OPT (R_{t'}) + \OPT(R^*_{\{u,v\}})$.
	Hence,
	for all collocations performed by \RM,
	and for final requests at $t=m$,
	\OPT pays at least 
	$\frac{1}{2}(
	\sum_{ t \in T } \OPT (R_t) +
	\sum_{\{u,v\}}\OPT (R^*_{\{u,v\}})
	) $.
	Then,
	the total cost to \OPT is

	\begin{align*} 	%\label{eq:costOPT}
		\OPT (\sigma)
		&=
		\frac{1}{2}
		\Big(
		\sum_{ t \in T} \OPT (R_t) 
		+ \sum_{\{u,v\}}\OPT (R^*_{\{u,v\}})
		\Big)	\\
		&\geq
		\Big(
		\sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
		\sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
		\Big)		
		=
		\frac{1}{2}		
		\Big(
		\sum_{ t \in T} \alpha  
		+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}
		\Big),
	\end{align*}
and
$
	\RM(\sigma)	/
	\OPT (\sigma)
	\leq
	2\Big(
	\sum_{ t \in T} 3\alpha +
	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
	\Big)	 \big /
	\Big(
	\sum_{ t \in T} \alpha  
	+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}  
	\Big) 	\leq 6.
$
\end{proof}

@@

\begin{algorithm}
	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
	\renewcommand{\algorithmicensure}{\textbf{Output:}}
	\begin{algorithmic}
		%        \Require 
		%        $k, \ell$,
		%        initial partition $P_I$,
		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
		%        \Ensure a~final partition $P_F$ 
		\STATE {For each node $v$ create a~singleton component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
		%\STATE{$P := P_I$}
		\label{line:initcomponents}
		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
		\STATE{Let $C_1 \ni u$ and $C_2 \ni v$ be the  components containing $u$ and $v$, respectively.}
		\IF{$C_1 \neq C_2$}
		\STATE {Merge $C_1$ and $C_2$ into one component $C'$ and
			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
		\IF{$C_1$ and $C_2$ are not collocated}
		\STATE {Move to a~partition closest to $P_I$ and respecting all components in $\mathcal{C}$.} 
		%		\COMMENT{move to a~partition closest to $P_I$}
		\label{line:rebalance} 
		\ENDIF
		\ENDIF
		\ENDFOR
	\end{algorithmic}
	\caption{Perfect Partition Learner (\PPL)}
	\label{alg:ppl}
\end{algorithm}
